---
title: "Tutorial on st_join with st_contains"
author: "John Little"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

## Load Library Packages

```{r library packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(mapview)
library(leafpop)
library(tigris)
library(DT)
```


## Load Data

```{r}
starbucksNC <- read_csv("https://raw.githubusercontent.com/libjohn/mapping-with-R/master/data/All_Starbucks_Locations_in_the_US_-_Map.csv") %>% 
  filter(State == "NC")

options(tigris_use_cache = TRUE)

durham_blocks <- tracts(state = "NC", county = "Durham", class = "sf", year = 2012)

durham_blocks <-  durham_blocks %>%  
  select(-6, -7, -8, -11, -12)

glimpse(durham_blocks)
```


## Wrangle

```{r}
sbux_drm <- starbucksNC %>% 
  filter(City == "Durham") %>% 
  select(`Store Number`, `Ownership Type`,
         `Phone Number`, City, Zip, Longitude,
         Latitude)


# assign my coordinates to a degree system
# documenting my XY values in degrees
# use, for example 4326 (best for US) ; 4269 (for US Agency)
sbux_drm_sf <- st_as_sf(sbux_drm,
                        coords = c("Longitude", "Latitude"),
                        crs = 4326)
st_crs(starbucksNC)
st_crs(sbux_drm_sf)
st_crs(durham_blocks)
# https://guides.library.duke.edu/r-geospatial/CRS
#       >    Projections / CRS > ESPG codes for CR's
```

## Tranform to a Projected CRS

Use one of the NC Projections

- State Plane North Carolina Feet	Lambert Conformal Conic	NAD83	feet	2264
- State Plane North Carolina Meters	Lambert Conformal Conic	NAD83	meters	32119
- UTM zone 17N [good for NC]	Transverse Mercator	WGS84	meters	32617

```{r}
# Transform to a Projection
durham_blocks <- st_transform(durham_blocks,
                              st_crs(2264))

sbux_drm_sf <- st_transform(sbux_drm_sf,
                            st_crs(durham_blocks)) 



st_crs(sbux_drm_sf)
st_crs(durham_blocks)

```

## First Lookie

```{r}
mapview(durham_blocks)
mapview(sbux_drm_sf)
mapview(list(durham_blocks, sbux_drm_sf))
```

## join

```{r}
shp_with_bux <- st_join(durham_blocks, sbux_drm_sf, join = st_contains, left = FALSE)

bux_within_shp <- st_join(sbux_drm_sf, durham_blocks, left = TRUE, join = st_within)
```


```{r}
plot(st_geometry(shp_with_bux))
plot(st_geometry(bux_within_shp))
```


```{r}
mapview(durham_blocks, col.regions = "grey",
        legend = FALSE,
        popup = popupTable(st_drop_geometry(durham_blocks), 
                           zcol = c(4,5),
                           row.numbers = FALSE,
                           feature.id = FALSE)) +
  mapview(shp_with_bux, 
          zcol = "NAME",
          layer.name = "Tract Name") +
  mapview(bux_within_shp,
          legend = FALSE,
          cex = 0,
          #cex = "AWATER",
          lwd = 3,
          alpha = .5,
          color = "red",
          zcol = "Store Number",
          popup =
            popupTable(st_drop_geometry(bux_within_shp),
                             zcol = c(1,2,3,5,10,11),
                           row.numbers = FALSE,
                           feature.id = FALSE))
```

## Display Table

```{r}
datatable(st_drop_geometry(bux_within_shp))
```


```{r}
bux_dt <- st_drop_geometry(bux_within_shp) %>% 
  select(1, GEOID, TractName = NAME, Ownership = `Ownership Type`) %>% 
  group_by(TractName) %>% 
  summarise(TotalStores = n()) %>% 
  arrange(-TotalStores) 

datatable(bux_dt)
```

# Choropleth 

Shade Tract shape by number of Stores in the Tract

```{r}
bux_tot_in_shape <- left_join(shp_with_bux, bux_dt, by = c("NAME" = "TractName"))

mapview(durham_blocks, col.regions = "grey",
        legend = FALSE,
        popup = popupTable(st_drop_geometry(durham_blocks), 
                           zcol = c(4,5),
                           row.numbers = FALSE,
                           feature.id = FALSE)) +
  mapview(bux_tot_in_shape, 
          zcol = "TotalStores",
          layer.name = "Total Stores by Tract",
          col.regions = c("royalblue1", "royalblue4")) +
  mapview(bux_within_shp,
          legend = FALSE,
          cex = 0,
          #cex = "AWATER",
          lwd = 4,
          color = "yellow",
          zcol = "Store Number",
          popup =
            popupTable(st_drop_geometry(bux_within_shp),
                             zcol = c(1,2,3,5,10,11),
                           row.numbers = FALSE,
                           feature.id = FALSE))

```

