---
title: "Hands On with counties st_join with st_contains"
author: "John Little"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

## Load Library Packages

```{r library packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(sf)
library(mapview)
library(leafpop)
library(tigris)
library(DT)
```


## Load Data

```{r}
starbucksNC <- read_csv("data/raw/All_Starbucks_Locations_in_the_US_-_Map.csv") %>% 
  filter(State == "NC")

options(tigris_use_cache = TRUE)

county_shapes <- counties(state = "NC", cb = TRUE, class = "sf", year = 2014) %>% 
  filter(NAME == "Durham" | 
           NAME == "Orange" | 
           NAME == "Chatham")
```


#### garbage chunk

```
county_names <- 
  read_excel("data/raw/County_Rural_Lookup.xlsx", 
             skip = 3) %>% 
  rename(countyfp = `2015 GEOID`) %>% 
  rename(county_name = `2015 Geography Name`) %>% 
  filter(State == "NC") %>% 
  mutate(county_name = str_extract(county_name, ".*,")) %>%
  mutate(county_name = str_replace(county_name, " County,", "")) %>% 
  mutate(countyfp = str_sub(countyfp, 3L)) %>% 
  select(countyfp, county_name) 

#durham_blocks <-  durham_blocks %>%  
#  select(-6, -7, -8, -11, -12)
```


## Wrangle

```{r}
starbucksNC <- starbucksNC %>% 
  #filter(City == "Durham") %>% 
  select(`Store Number`, `Ownership Type`,
         `Phone Number`, City, Zip, Longitude,
         Latitude)


# assign my coordinates to a degree system
# documenting my XY values in degrees
# use, for example 4326 (best for US) ; 4269 (for US Agency)
starbucksNC_sf <- st_as_sf(starbucksNC,
                        coords = c("Longitude", "Latitude"),
                        crs = 4326)
```


```{r}
glimpse(county_shapes)
```



```{r}
plot(county_shapes)
mapview(county_shapes)
```



## Tranform to a Projected CRS

Use one of the NC Projections

- State Plane North Carolina Feet	Lambert Conformal Conic	NAD83	feet	2264
- State Plane North Carolina Meters	Lambert Conformal Conic	NAD83	meters	32119
- UTM zone 17N [good for NC]	Transverse Mercator	WGS84	meters	32617

### Useful CRS background

- `st_crs(starbucksNC)`  # get the assigned CRS

- [Mark's Handy EPSG table](https://guides.library.duke.edu/r-geospatial/CRS#s-lib-ctab-21326020-2)




```{r}
# Transform to a Projection
county_shapes <- st_transform(county_shapes,
                              st_crs(2264))

starbucksNC_sf <- st_transform(starbucksNC_sf,
                            st_crs(county_shapes)) 



st_crs(county_shapes)
st_crs(starbucksNC_sf)

```

## First Lookie

```{r}
mapview(county_shapes)
mapview(starbucksNC_sf)
mapview(list(county_shapes, starbucksNC_sf),
        color = list("red", "blue"))
```

## join

```{r}
shp_with_bux <- st_join(county_shapes, starbucksNC_sf, join = st_contains, left = FALSE)

#bux_within_shp <- st_join(starbucksNC_sf, county_shapes, left = TRUE, join = st_within)

bux_within_shp <- st_join(starbucksNC_sf, county_shapes, left = FALSE, join = st_within)

```


```{r}
plot(st_geometry(shp_with_bux))
plot(st_geometry(bux_within_shp))
#plot(st_geometry(foo))
```


```{r}
#mapview(county_shapes, col.regions = "grey",
#        legend = FALSE,
#        popup = popupTable(st_drop_geometry(county_shapes), 
 #                          zcol = c(4,5),
#                           row.numbers = FALSE,
#                           feature.id = FALSE)) +
mapview(shp_with_bux, 
          zcol = "NAME",
          layer.name = "NAME",
                    popup =
            popupTable(st_drop_geometry(bux_within_shp),
                           zcol = c("Store Number", 
                                    "Ownership Type", 
                                    "Phone Number", 
                                    "City", "NAME","Zip", 
                                    "GEOID", "ALAND",
                                    "AWATER"),
                           row.numbers = TRUE,
                           feature.id = FALSE)) +
  mapview(bux_within_shp,
          legend = FALSE,
          cex = 0,
          #cex = "AWATER",
          lwd = 3,
          alpha = .5,
          color = "red",
          zcol = "Store Number",
          popup =
            popupTable(st_drop_geometry(bux_within_shp),
                             zcol = c(1,2,3,5,10,11),
                           row.numbers = FALSE,
                           feature.id = FALSE))
```

## Display Table

```{r}
datatable(st_drop_geometry(bux_within_shp))
```


```{r}
bux_dt <- st_drop_geometry(bux_within_shp) %>% 
  select(1, GEOID, TractName = NAME, Ownership = `Ownership Type`) %>% 
  group_by(TractName) %>% 
  summarise(TotalStores = n()) %>% 
  arrange(-TotalStores) 

datatable(bux_dt)
```

# Choropleth 

Shade Tract shape by number of Stores in the Tract

```{r}
bux_tot_in_shape <- left_join(shp_with_bux, bux_dt, by = c("NAME" = "TractName"))

mapview(bux_tot_in_shape, 
          zcol = "TotalStores",
          layer.name = "Total Stores by Tract",
          col.regions = c("royalblue", "royalblue2", "royalblue4")) +
  mapview(bux_within_shp,
          legend = FALSE,
          cex = 0,
          #cex = "AWATER",
          lwd = 4,
          color = "yellow",
          zcol = "Store Number",
          popup =
            popupTable(st_drop_geometry(bux_within_shp),
                             zcol = c(1,2,3,5,10,11),
                           row.numbers = FALSE,
                           feature.id = FALSE))

```





```{r}
bux_within_shp %>% 
  ggplot() +
  geom_sf()


bux_tot_in_shape %>% 
  mutate(TotalStores = as.character(TotalStores)) %>%
  ggplot() +
  geom_sf(aes(fill = fct_infreq(TotalStores))) +
  scale_fill_brewer("Total Stores", type = "seq", direction = -1, palette = "Greens") +
  geom_sf(data = bux_within_shp, col = "purple") 

  
```



